---
title: "diet analysis"
author: "Dani Blumstein"
date: "9/30/2020"
output: html_document
---
libs
```{r}
library(ggplot2)
library(tidyverse)
library(car) 
library(lubridate)
library(ggpubr)
library(ggpmisc)
library(gridExtra)
library(rlist)
library(broom)
library(nlme)
library(rstatix)
library(compositions)
library(maditr)
library(RColorBrewer)
library(changepoint)
library(mcp)
library(EnvCpt)
require(grDevices)
library(ggfortify)
library(tsbox)
library(zoo)
library(bcp)
library(plyr)
library(cowplot)
library(broom)
library(gridGraphics)
library(ggsignif)
library(ARTool)
library(dplyr)
```
function for saving images 
```{r}
library(reshape)
library(reshape2)
library(tidyverse)
library(viridis)
library(data.table)
library(gridExtra)
library(grid)


save <- function(name, fig_len, fig_width,plot_name)
{
library(Cairo)

Cairo::Cairo(
  fig_len, #length
  fig_width, #width
  file = paste(name, ".png", sep = ""),
  type = "png", #tiff
  bg = "transparent", #white or transparent depending on your requirement 
  dpi = 300,
  units = "cm" #you can change to pixels etc 
)
plot(plot_name) #p is your graph object 
dev.off()
}
```

read in the data
```{r}
electrolyte_data_diet <- read.csv("/Users/danielleblumstein/OneDrive - USNH/BOX/Cactus_Mouse_Physiology/data/5015_lowfat/electrolyte_data.csv")
setwd("/Users/danielleblumstein/OneDrive - USNH/BOX/Cactus_Mouse_Physiology/data/5015_lowfat")

data <- read_csv("diet_analysis_data.csv", 
                 col_types = cols(Sex = col_character(),
                                  EE = col_double(),
                                  H2Omg = col_double(),
                                  RQ = col_double(),
                                  Animal_ID = col_character(),
                                  Deg_C = col_double(),
                                  weight = col_double(),
                                  experiment = col_character(),
                                  StartTime = col_character(), #col_time(format = "%H:%M:%S"), - changed for easy use of lubridate
                                  SD_VCO2 = col_double(),
                                  SD_VO2 = col_double(),
                                  SD_H2Omg = col_double(),
                                  VO2 = col_double(),
                                  VCO2 = col_double(),
                                  StartDate = col_date(format = "%Y-%m-%d"), 
                                  hour = col_integer()))

```

Establish day and night
```{r}
data$time_in_S <- period_to_seconds(hms(data$StartTime)) #Total seconds to reach that point in the day

#Replace mis-typed weight (29.980) with correct weight (19.980)
data$weight[data$weight == 29.980] <- 19.980

# Establish when each interval/transition starts and stops
#Daytime interval: hrs:8:00-21:00
daytime_interval <- period_to_seconds(hms("09:00:00")):period_to_seconds(hms("20:00:00"))
#Night time: hrs 22:00-5:00 
nighttime_interval <- c((period_to_seconds(hms("21:00:01")):period_to_seconds(hms("24:59:59"))), #evening portion of 'nighttime'
                        (period_to_seconds(hms("00:00:00")):period_to_seconds(hms("06:00:00")))) #morning portion of 'nightitme'
#Morning transition (t1): 6:00-9:00
t1_interval <- period_to_seconds(hms("06:00:00")):period_to_seconds(hms("09:00:00"))
#Evening transition (t2): 20-21:00
t2_interval <- period_to_seconds(hms("20:00:00")):period_to_seconds(hms("21:00:00"))

#### Create data subsets for males and females for each experiment (day, night, baseline [BL]) and each time block [e.g., t1/transition1, daytime, t2/transistion2, nighttime])
# all males and females across all 3 experiments (BL, hot, cold)
all_M = data[data$Sex == 'M', ]
all_F = data[data$Sex == 'F', ]
```

# IDENTIFY AND REMOVE OUTLIERS
### MALES ###
```{r}
count = 1
dependent_variables = c("EE", "RQ", "VO2", "VCO2", "H2Omg")
varname_list = c("EE_OLs","RQ_OLs","VO2_OLs","VCO2_OLs","H2O_OLs")

for (DV in dependent_variables){
  print(DV)
    model <- lm(as.formula(paste0(DV, " ~ weight + experiment")), data = all_M)
  model.metrics <- augment(model) %>% select(-.hat, -.sigma, -.fitted)
  summ <- model.metrics %>% filter(abs(.std.resid) > 3) %>% as.data.frame()

    OL_list = c()
  masterlist = c()
  for (each_outlier_row in 1:nrow(summ)){
    this_weight <- summ[each_outlier_row, "weight"]
    this_DV  <- summ[each_outlier_row, DV]
    OL_list <- c(OL_list, (which(all_M$weight == this_weight & all_M[DV] == this_DV)))

  }
  print(OL_list)
  #Assign list of outliers to specified variable lists (OL_list_EE, RQ, VO2, VCO2, mgH2O)
  assign(paste("OL_list_", DV, sep = ""), OL_list)  
}

masterlist <- c(OL_list_EE, OL_list_RQ, OL_list_VO2, OL_list_VCO2, OL_list_H2Omg)
masterlist_noDup <- unique(masterlist)

all_noOL_M <- all_M[-c(masterlist_noDup),]
dim(all_M) #6470 
dim(all_noOL_M) #6255
write.csv(all_noOL_M, "all_noOL_M.csv", row.names = FALSE)
```

# IDENTIFY AND REMOVE OUTLIERS
#### FEMALES ####
```{r}
count = 1
dependent_variables = c("EE", "RQ", "VO2", "VCO2", "H2Omg")
varname_list = c("EE_OLs","RQ_OLs","VO2_OLs","VCO2_OLs","H2O_OLs")

for (DV in dependent_variables){
  print(DV)
  model <- lm(as.formula(paste0(DV, " ~ weight + experiment")), data = all_F)
  model.metrics <- augment(model) %>% select(-.hat, -.sigma, -.fitted)
  summ <- model.metrics %>% filter(abs(.std.resid) > 3) %>% as.data.frame()
  OL_list = c()
  masterlist = c()
  for (each_outlier_row in 1:nrow(summ)){
    this_weight <- summ[each_outlier_row, "weight"]
    this_DV  <- summ[each_outlier_row, DV]
    OL_list <- c(OL_list, (which(all_F$weight == this_weight & all_F[DV] == this_DV)))
    
  }
  print(OL_list)
  assign(paste("OL_list_", DV, sep = ""), OL_list)  
}

masterlist <- c(OL_list_EE, OL_list_RQ, OL_list_VO2, OL_list_VCO2, OL_list_H2Omg)
masterlist_noDup <- unique(masterlist)
all_noOL_F <- all_F[-c(masterlist_noDup),]
write.csv(all_noOL_F, "all_noOL_F.csv", row.names=FALSE)
```

datasets and subset for BL and LF
```{r}
all_noOL_F
all_noOL_M

all_day_noOL_M = all_noOL_M[all_noOL_M$time_in_S %in% daytime_interval, ]
all_day_noOL_M$interval <- "light"
  
all_night_noOL_M = all_noOL_M[all_noOL_M$time_in_S %in% nighttime_interval, ]
all_night_noOL_M$interval <- "dark"

all_t1_noOL_M = all_noOL_M[all_noOL_M$time_in_S %in% t1_interval, ]
all_t1_noOL_M$interval <- "T1"

all_t2_noOL_M = all_noOL_M[all_noOL_M$time_in_S %in% t2_interval, ]
all_t2_noOL_M$interval <- "T2"

all_day_noOL_F = all_noOL_F[all_noOL_F$time_in_S %in% daytime_interval, ]
all_day_noOL_F$interval <- "light"

all_night_noOL_F = all_noOL_F[all_noOL_F$time_in_S %in% nighttime_interval, ]
all_night_noOL_F$interval <- "dark"

all_t1_noOL_F = all_noOL_F[all_noOL_F$time_in_S %in% t1_interval, ]
all_t1_noOL_F$interval <- "T1"

all_t2_noOL_F = all_noOL_F[all_noOL_F$time_in_S %in% t2_interval, ]
all_t2_noOL_F$interval <- "T2"

anov_data <- rbind(all_day_noOL_M,all_night_noOL_M,all_t1_noOL_M,all_t2_noOL_M,all_day_noOL_F,all_night_noOL_F,all_t1_noOL_F,all_t2_noOL_F)

all_data_m <- rbind(all_day_noOL_M,all_night_noOL_M,all_t1_noOL_M,all_t2_noOL_M)
all_data_f <- rbind(all_day_noOL_F,all_night_noOL_F,all_t1_noOL_F,all_t2_noOL_F)
```

mgH2O
```{r}
#area under the curve
library(DescTools)
library(emmeans)

##H2O
by_id <- list()
by_id_inter <- list()

data <- c()
final <- data.frame()

#sintegral(xfx)

 for (ID in unique(anov_data$Animal_ID)) {
      for (inter in unique(anov_data$interval)){ 
        by_id[[ID]] <- filter(anov_data, Animal_ID == ID)
        by_id_inter[[paste(ID, inter, sep = "_")]] <- filter(by_id[[ID]], interval == inter)
        weight <- unique(by_id_inter[[paste(ID,  inter, sep = "_")]]$weight)
        sex <- unique(by_id_inter[[paste(ID, inter, sep = "_")]]$Sex)
        exp <- unique(by_id_inter[[paste(ID, inter, sep = "_")]]$experiment)
        z <- AUC(by_id_inter[[paste(ID, inter, sep = "_")]]$H2Omg, by_id_inter[[paste(ID, inter, sep = "_")]]$time_in_S)
        data <- c(ID,inter,exp, z, weight, sex)
        final <- rbind(final, data)
    }
  }

auc_data <- final %>% 
  rename(
    Animal_ID = X.1079.,
    diet = X.Low.Fat.,
    interval = X.light.,
    mgH2O = X.11470.63701685.,
    weight = X.25.506.,
    sex = X.M.
    )

auc_data$Animal_ID <- as.factor(auc_data$Animal_ID)
auc_data$diet <- as.factor(auc_data$diet)
auc_data$interval <- as.factor(auc_data$interval)
auc_data$mgH2O <- as.numeric(auc_data$mgH2O)
auc_data$weight <- as.numeric(auc_data$weight)
auc_data$sex<- as.factor(auc_data$sex)
auc_data$trt<- as.factor(paste(auc_data$sex, auc_data$diet))

ggplot(auc_data, aes(x=mgH2O)) + 
  geom_density()

ylim <- max(auc_data$mgH2O)-3000
ggplot(auc_data, aes(x = diet, y = mgH2O)) +
  geom_boxplot(size = .75) +
  geom_jitter(alpha = .5) +
  stat_compare_means(aes(group=diet), label = "p.signif", method="t.test", label.y = ylim)+
  facet_grid(sex ~ interval, margins = TRUE) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))+
  labs(x = "", y = "RWL")

ggplot(auc_data, aes(x=interval, y=mgH2O, color = trt))+
  geom_boxplot()

#################
#time intervals
#DAY
day <- filter(auc_data, interval == "light")
day$mgH2O <- as.numeric(day$mgH2O)

#The ANOVA
mod<-lm(mgH2O ~ diet * sex , day)
anova(mod)

#anova with x in covar as response
anovaX_mod<-lm(weight ~ diet * sex, day)
anova(anovaX_mod)

#transform
day$trans <- log10(10*day$mgH2O)
mod<-lm(trans ~ diet * sex, day)
anova(mod)

#TESTING ASSUMPTIONS
#testing for homo of slopes
#This interaction must be NS in order for us to carry out the ANCOVA.
slopes_mod<-lm(trans ~ sex + diet + weight + diet:weight, day)
anova(slopes_mod)

#Perform a Shapiro-Wilk test for normality of residuals
day$resids <- residuals(mod)
shapiro.test(day$resids)

#Perform Levene's Test for homogenity of variances
#library(car)
leveneTest(trans ~ diet, data = day)

#Perform a Tukey 1-df Test for Non-additivity
day$preds <- predict(mod)
day$sq_preds <- day$preds^2
tukey_mod<-lm(trans ~ sex + diet + sq_preds, day)
anova(tukey_mod)

#adjusting for differences in weight
ancova_mod<-lm(trans ~ diet + sex + weight, day)
Anova(ancova_mod, type = 2)
#to get beta val
summary(ancova_mod)

xbar <- mean(day$weight)
day$z <- day$trans +-0.0005936*(day$weight - xbar)

anovaZ_mod<-lm(z ~ diet * sex, day)
anova(anovaZ_mod)

#post adj assumption testing
#norm of residuals 
day$anovaZ_resids <- residuals(anovaZ_mod)
shapiro.test(day$anovaZ_resids)

#homo of variety var
leveneTest(z ~ sex, data = day)

#Test for Nonadditivity
day$anovaZ_preds <- predict(anovaZ_mod)
day$sq_anovaZ_preds <- day$anovaZ_preds^2
tukeyZ_mod<-lm(z ~ sex + diet + sq_anovaZ_preds, day)
anova(tukeyZ_mod)

emmip(anovaZ_mod, sex~diet, CIs=TRUE)

library(emmeans)
lsm <- emmeans(anovaZ_mod, specs = pairwise ~ sex:diet)
lsm$contrasts

#################
#NIGHT
night <- filter(auc_data, interval == "dark")
night$mgH2O <- as.numeric(night$mgH2O)

#The ANOVA
mod<-lm(mgH2O ~ diet * sex , night)
anova(mod)

#transform
night$trans <- log10(10*night$mgH2O)
mod<-lm(trans ~ diet * sex, night)
anova(mod)

#anova with x in covar as response
anovaX_mod<-lm(weight ~ diet * sex, night)
anova(anovaX_mod)

#TESTING ASSUMPTIONS
#testing for homo of slopes
#This interaction must be NS in order for us to carry out the ANCOVA.
slopes_mod<-lm(trans ~ sex + diet + weight + diet:weight, night)
anova(slopes_mod)

#failed without transformation...passed with transformation
#Perform a Shapiro-Wilk test for normality of residuals
night$resids <- residuals(mod)
shapiro.test(night$resids)

#Perform Levene's Test for homogenity of variances
#library(car)
leveneTest(trans ~ diet, data = night)

#Perform a Tukey 1-df Test for Non-additivity
night$preds <- predict(mod)
night$sq_preds <- night$preds^2
tukey_mod<-lm(trans ~ sex + diet + sq_preds, night)
anova(tukey_mod)


#adjusting for differences in weight
ancova_mod<-lm(trans ~ diet + sex + weight, night)
Anova(ancova_mod, type = 2)
#to get beta val
summary(ancova_mod)

xbar <- mean(night$weight)
night$z <- night$trans + 0.0003139*(night$weight - xbar)

anovaZ_mod<-lm(z ~ diet * sex, night)
anova(anovaZ_mod)

#post adj assumption testing
#norm of residuals 
night$anovaZ_resids <- residuals(anovaZ_mod)
shapiro.test(night$anovaZ_resids)

#homo of variety var
leveneTest(z ~ diet, data = night)

#Test for Nonadditivity
night$anovaZ_preds <- predict(anovaZ_mod)
night$sq_anovaZ_preds <- night$anovaZ_preds^2
tukeyZ_mod<-lm(z ~ sex + diet + sq_anovaZ_preds, night)
anova(tukeyZ_mod)


library(emmeans)
lsm <- emmeans(anovaZ_mod, specs = pairwise ~ sex:diet)
lsm$contrast

#################
#T1
t1 <- filter(auc_data, interval == "T1")
t1$mgH2O <- as.numeric(t1$mgH2O)

#The ANOVA
mod<-lm(mgH2O ~ diet * sex , t1)
anova(mod)

#transform
t1$trans <- log10(10*t1$mgH2O)
mod<-lm(trans ~ diet * sex, t1)
anova(mod)

#anova with x in covar as response
anovaX_mod<-lm(weight ~ diet * sex, t1)
anova(anovaX_mod)

#TESTING ASSUMPTIONS
#testing for homo of slopes
#This interaction must be NS in order for us to carry out the ANCOVA.
slopes_mod<-lm(trans ~ sex + diet + weight + diet:weight, t1)
anova(slopes_mod)

#fails without transformations
#Perform a Shapiro-Wilk test for normality of residuals
t1$resids <- residuals(mod)
shapiro.test(t1$resids)

#Perform Levene's Test for homogenity of variances
#library(car)
leveneTest(trans ~ diet, data = t1)

#Perform a Tukey 1-df Test for Non-additivity
t1$preds <- predict(mod)
t1$sq_preds <- t1$preds^2
tukey_mod<-lm(trans ~ sex + diet + sq_preds, t1)
anova(tukey_mod)

#adjusting for differences in weight
ancova_mod<-lm(trans ~ diet + sex + weight, t1)
Anova(ancova_mod, type = 2)
#to get beta val
summary(ancova_mod)

xbar <- mean(t1$weight)
t1$z <- t1$trans + -0.006752*(t1$weight - xbar)

anovaZ_mod<-lm(z ~ diet * sex, t1)
anova(anovaZ_mod)

#post adj assumption testing
#norm of residuals 
t1$anovaZ_resids <- residuals(anovaZ_mod)
shapiro.test(t1$anovaZ_resids)

#homo of variety var
leveneTest(z ~ sex, data = t1)

#Test for Nonadditivity
t1$anovaZ_preds <- predict(anovaZ_mod)
t1$sq_anovaZ_preds <- t1$anovaZ_preds^2
tukeyZ_mod<-lm(z ~ sex + diet + sq_anovaZ_preds, t1)
anova(tukeyZ_mod)

library(emmeans)
lsm <- emmeans(anovaZ_mod, specs = pairwise ~ sex:diet)
lsm$contrasts


#################
#T2
t2 <- filter(auc_data, interval == "T2")
t2$mgH2O <- as.numeric(t2$mgH2O)

#The ANOVA
mod<-lm(mgH2O ~ diet * sex , t2)
anova(mod)

#anova with x in covar as response
anovaX_mod<-lm(weight ~ diet * sex, t2)
anova(anovaX_mod)

#TESTING ASSUMPTIONS
#testing for homo of slopes
#This interaction must be NS in order for us to carry out the ANCOVA.
slopes_mod<-lm(mgH2O ~ sex + diet + weight + diet:weight, t2)
anova(slopes_mod)

#Perform a Shapiro-Wilk test for normality of residuals
t2$resids <- residuals(mod)
shapiro.test(t2$resids)

#Perform Levene's Test for homogenity of variances
#library(car)
leveneTest(mgH2O ~ diet, data = t2)

#Perform a Tukey 1-df Test for Non-additivity
t2$preds <- predict(mod)
t2$sq_preds <- t2$preds^2
tukey_mod<-lm(mgH2O ~ sex + diet + sq_preds, t2)
anova(tukey_mod)

#adjusting for differences in weight
ancova_mod<-lm(mgH2O ~ diet + sex + weight, t2)
Anova(ancova_mod, type = 2)
#to get beta val
summary(ancova_mod)

xbar <- mean(t2$weight)
t2$z <- t2$mgH2O + -361.8*(t2$weight - xbar)

anovaZ_mod<-lm(z ~ diet * sex, t2)
anova(anovaZ_mod)

#post adj assumption testing
#norm of residuals 
t2$anovaZ_resids <- residuals(anovaZ_mod)
shapiro.test(t2$anovaZ_resids)

#homo of variety var
leveneTest(z ~ sex, data = t2)

#Test for Nonadditivity
t2$anovaZ_preds <- predict(anovaZ_mod)
t2$sq_anovaZ_preds <- t2$anovaZ_preds^2
tukeyZ_mod<-lm(z ~ sex + diet + sq_anovaZ_preds, t2)
anova(tukeyZ_mod)

library(emmeans)
lsm <- emmeans(anovaZ_mod, specs = pairwise ~ sex:diet)
lsm$contrasts

auc_master <- data.frame(auc_data$Animal_ID, auc_data$interval, auc_data$diet, auc_data$sex, auc_data$mgH2O)
```

EE
```{r}
by_id <- list()
by_id_inter <- list()

data <- c()
final <- data.frame()

 for (ID in unique(anov_data$Animal_ID)) {
      for (inter in unique(anov_data$interval)){ 
        by_id[[ID]] <- filter(anov_data, Animal_ID == ID)
        by_id_inter[[paste(ID, inter, sep = "_")]] <- filter(by_id[[ID]], interval == inter)
        weight <- unique(by_id_inter[[paste(ID,  inter, sep = "_")]]$weight)
        sex <- unique(by_id_inter[[paste(ID, inter, sep = "_")]]$Sex)
        exp <- unique(by_id_inter[[paste(ID, inter, sep = "_")]]$experiment)
        z <- AUC(by_id_inter[[paste(ID, inter, sep = "_")]]$EE, by_id_inter[[paste(ID, inter, sep = "_")]]$time_in_S)
        norm_z <- z/(max(by_id_inter[[paste(ID, inter, sep = "_")]]$time_in_S)-min(by_id_inter[[paste(ID, inter, sep = "_")]]$time_in_S))
        data <- c(ID,inter,exp, z,norm_z, weight, sex)
        final <- rbind(final, data)
    }
  }


auc_data <- final %>% 
  rename(
    Animal_ID = X.1079.,
    diet = X.Low.Fat.,
    interval = X.light.,
    EE =X.12993.9304744685.,
    weight = X.25.506.,
    sex = X.M.
    )

auc_data$Animal_ID <- as.factor(auc_data$Animal_ID)
auc_data$diet <- as.factor(auc_data$diet)
auc_data$interval <- as.factor(auc_data$interval)
auc_data$EE <- as.numeric(auc_data$EE)
auc_data$weight <- as.numeric(auc_data$weight)
auc_data$sex<- as.factor(auc_data$sex)
auc_data$trt<- as.factor(paste(auc_data$sex, auc_data$diet))

ggplot(auc_data, aes(x=EE)) + 
  geom_density()

ylim <- max(auc_data$EE)-3000
ggplot(auc_data, aes(x = diet, y = EE)) +
  geom_boxplot(size = .75) +
  geom_jitter(alpha = .5) +
  stat_compare_means(aes(group=diet), label = "p.signif", method="t.test", label.y = ylim)+
  facet_grid(sex ~ interval, margins = TRUE) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))+
  labs(x = "", y = "RWL")

ggplot(auc_data, aes(x=interval, y=EE, color = trt))+
  geom_boxplot()

#################
#time intervals
#DAY
day <- filter(auc_data, interval == "light")
day$EE <- as.numeric(day$EE)

#The ANOVA
mod<-lm(EE ~ diet * sex , day)
anova(mod)

#anova with x in covar as response
anovaX_mod<-lm(weight ~ diet * sex, day)
anova(anovaX_mod)

#TESTING ASSUMPTIONS
#testing for homo of slopes
#This interaction must be NS in order for us to carry out the ANCOVA.
slopes_mod<-lm(EE ~ sex + diet + weight + sex:weight, day)
anova(slopes_mod)

#Perform a Shapiro-Wilk test for normality of residuals
day$resids <- residuals(mod)
shapiro.test(day$resids)

#Perform Levene's Test for homogenity of variances
#library(car)
leveneTest(EE ~ sex, data = day)

#Perform a Tukey 1-df Test for Non-additivity
day$preds <- predict(mod)
day$sq_preds <- day$preds^2
tukey_mod<-lm(EE ~ sex + diet + sq_preds, day)
anova(tukey_mod)

#adjusting for differences in weight
ancova_mod<-lm(EE ~ diet + sex + weight, day)
Anova(ancova_mod, type = 2)
#to get beta val
summary(ancova_mod)

xbar <- mean(day$weight)
day$z <- day$EE + -177.3*(day$weight - xbar)

anovaZ_mod<-lm(z ~ diet * sex, day)
anova(anovaZ_mod)

#post adj assumption testing
#norm of residuals 
day$anovaZ_resids <- residuals(anovaZ_mod)
shapiro.test(day$anovaZ_resids)

#homo of variety var
leveneTest(z ~ sex, data = day)

#Test for Nonadditivity
day$anovaZ_preds <- predict(anovaZ_mod)
day$sq_anovaZ_preds <- day$anovaZ_preds^2
tukeyZ_mod<-lm(z ~ sex + diet + sq_anovaZ_preds, day)
anova(tukeyZ_mod)

library(emmeans)
lsm <- emmeans(anovaZ_mod, specs = pairwise ~ sex:diet)
lsm$contrasts

#################
#NIGHT
night <- filter(auc_data, interval == "dark")
night$EE <- as.numeric(night$EE)

#The ANOVA
mod<-lm(EE ~ diet * sex , night)
anova(mod)

#transform
night$trans <- log10(10*night$EE)
mod<-lm(trans ~ diet * sex, night)
anova(mod)

#anova with x in covar as response
anovaX_mod<-lm(weight ~ diet * sex, night)
anova(anovaX_mod)

#TESTING ASSUMPTIONS
#testing for homo of slopes
#This interaction must be NS in order for us to carry out the ANCOVA.
slopes_mod<-lm(trans ~ sex + diet + weight + diet:weight, night)
anova(slopes_mod)

#failed without transformation...passed with transformation
#Perform a Shapiro-Wilk test for normality of residuals
night$resids <- residuals(mod)
shapiro.test(night$resids)

#Perform Levene's Test for homogenity of variances
#library(car)
leveneTest(trans ~ diet, data = night)

#Perform a Tukey 1-df Test for Non-additivity
night$preds <- predict(mod)
night$sq_preds <- night$preds^2
tukey_mod<-lm(trans ~ sex + diet + sq_preds, night)
anova(tukey_mod)


#adjusting for differences in weight
ancova_mod<-lm(trans ~ diet + sex + weight, night)
Anova(ancova_mod, type = 2)
#to get beta val
summary(ancova_mod)

xbar <- mean(night$weight)
night$z <- night$trans + 0.008468*(night$weight - xbar)

anovaZ_mod<-lm(z ~ diet * sex, night)
anova(anovaZ_mod)

#post adj assumption testing
#norm of residuals 
night$anovaZ_resids <- residuals(anovaZ_mod)
shapiro.test(night$anovaZ_resids)

#homo of variety var
leveneTest(z ~ diet, data = night)

#Test for Nonadditivity
night$anovaZ_preds <- predict(anovaZ_mod)
night$sq_anovaZ_preds <- night$anovaZ_preds^2
tukeyZ_mod<-lm(z ~ sex + diet + sq_anovaZ_preds, night)
anova(tukeyZ_mod)


library(emmeans)
lsm <- emmeans(anovaZ_mod, specs = pairwise ~ sex:diet)
lsm$contrast

#################
#T1
t1 <- filter(auc_data, interval == "T1")
t1$EE <- as.numeric(t1$EE)

#The ANOVA
mod<-lm(EE ~ diet * sex , t1)
anova(mod)

#anova with x in covar as response
anovaX_mod<-lm(weight ~ diet * sex, t1)
anova(anovaX_mod)

#TESTING ASSUMPTIONS
#testing for homo of slopes
#This interaction must be NS in order for us to carry out the ANCOVA.
slopes_mod<-lm(EE ~ sex + diet + weight + diet:weight, t1)
anova(slopes_mod)

#Perform a Shapiro-Wilk test for normality of residuals
t1$resids <- residuals(mod)
shapiro.test(t1$resids)

#Perform Levene's Test for homogenity of variances
#library(car)
leveneTest(EE ~ diet, data = t1)

#Perform a Tukey 1-df Test for Non-additivity
t1$preds <- predict(mod)
t1$sq_preds <- t1$preds^2
tukey_mod<-lm(EE ~ sex + diet + sq_preds, t1)
anova(tukey_mod)

#adjusting for differences in weight
ancova_mod<-lm(EE ~ diet + sex + weight, t1)
Anova(ancova_mod, type = 2)
#to get beta val
summary(ancova_mod)

xbar <- mean(t1$weight)
t1$z <- t1$EE + 31.84*(t1$weight - xbar)

anovaZ_mod<-lm(z ~ diet * sex, t1)
anova(anovaZ_mod)

#post adj assumption testing
#norm of residuals 
t1$anovaZ_resids <- residuals(anovaZ_mod)
shapiro.test(t1$anovaZ_resids)

#homo of variety var
leveneTest(z ~ sex, data = t1)

#Test for Nonadditivity
t1$anovaZ_preds <- predict(anovaZ_mod)
t1$sq_anovaZ_preds <- t1$anovaZ_preds^2
tukeyZ_mod<-lm(z ~ sex + diet + sq_anovaZ_preds, t1)
anova(tukeyZ_mod)

library(emmeans)
lsm <- emmeans(anovaZ_mod, specs = pairwise ~ sex:diet)
lsm$contrasts

#################
#T2
t2 <- filter(auc_data, interval == "T2")
t2$EE <- as.numeric(t2$EE)

#The ANOVA
mod<-lm(EE ~ diet * sex , t2)
anova(mod)

#transform
t2$trans <- log10(10*t2$EE)
mod<-lm(trans ~ diet * sex, t2)
anova(mod)

#anova with x in covar as response
anovaX_mod<-lm(weight ~ diet * sex, t2)
anova(anovaX_mod)

#TESTING ASSUMPTIONS
#testing for homo of slopes
#This interaction must be NS in order for us to carry out the ANCOVA.
slopes_mod<-lm(trans ~ sex + diet + weight + diet:weight, t2)
anova(slopes_mod)

#fails without transformations
#Perform a Shapiro-Wilk test for normality of residuals
t2$resids <- residuals(mod)
shapiro.test(t2$resids)

#Perform Levene's Test for homogenity of variances
#library(car)
leveneTest(trans ~ diet, data = t2)

#Perform a Tukey 1-df Test for Non-additivity
t2$preds <- predict(mod)
t2$sq_preds <- t2$preds^2
tukey_mod<-lm(trans ~ sex + diet + sq_preds, t2)
anova(tukey_mod)

#adjusting for differences in weight
ancova_mod<-lm(trans ~ diet + sex + weight, t2)
Anova(ancova_mod, type = 2)
#to get beta val
summary(ancova_mod)

xbar <- mean(t2$weight)
t2$z <- t2$trans + 0.008275*(t2$weight - xbar)

anovaZ_mod<-lm(z ~ diet * sex, t2)
anova(anovaZ_mod)

#post adj assumption testing
#norm of residuals 
t2$anovaZ_resids <- residuals(anovaZ_mod)
shapiro.test(t2$anovaZ_resids)

#homo of variety var
leveneTest(z ~ sex, data = t2)

#Test for Nonadditivity
t2$anovaZ_preds <- predict(anovaZ_mod)
t2$sq_anovaZ_preds <- t2$anovaZ_preds^2
tukeyZ_mod<-lm(z ~ sex + diet + sq_anovaZ_preds, t2)
anova(tukeyZ_mod)

library(emmeans)
lsm <- emmeans(anovaZ_mod, specs = pairwise ~ sex:diet)
lsm$contrasts

auc_master <- auc_master %>% 
  rename(
    Animal_ID = auc_data.Animal_ID,
    diet = auc_data.diet,
    interval = auc_data.interval,
    sex = auc_data.sex
    )


auc_master <- merge(auc_master,auc_data,by=c("Animal_ID","interval", "diet", "sex"))
```

RQ 
```{r}
by_id <- list()
by_id_inter <- list()

data <- c()
final <- data.frame()

 for (ID in unique(anov_data$Animal_ID)) {
      for (inter in unique(anov_data$interval))
      { 
        by_id[[ID]] <- filter(anov_data, Animal_ID == ID)
        by_id_inter[[paste(ID, inter, sep = "_")]] <- filter(by_id[[ID]], interval == inter)
        weight <- unique(by_id_inter[[paste(ID,  inter, sep = "_")]]$weight)
        sex <- unique(by_id_inter[[paste(ID, inter, sep = "_")]]$Sex)
        exp <- unique(by_id_inter[[paste(ID, inter, sep = "_")]]$experiment)
        z <- AUC(by_id_inter[[paste(ID, inter, sep = "_")]]$RQ, by_id_inter[[paste(ID, inter, sep = "_")]]$time_in_S)
        norm_z <- z/(max(by_id_inter[[paste(ID, inter, sep = "_")]]$time_in_S)-min(by_id_inter[[paste(ID, inter, sep = "_")]]$time_in_S))
        data <- c(ID,inter,exp, z,norm_z, weight, sex)
        final <- rbind(final, data)
    }
  }


auc_data <- final %>% 
  rename(
    Animal_ID = X.1079.,
    diet = X.Low.Fat.,
    interval = X.light.,
    RQ =X.23889.4482013383.,
    weight = X.25.506.,
    sex = X.M.
    )

auc_data$Animal_ID <- as.factor(auc_data$Animal_ID)
auc_data$diet <- as.factor(auc_data$diet)
auc_data$interval <- as.factor(auc_data$interval)
auc_data$RQ <- as.numeric(auc_data$RQ)
auc_data$weight <- as.numeric(auc_data$weight)
auc_data$sex<- as.factor(auc_data$sex)
auc_data$trt<- as.factor(paste(auc_data$sex, auc_data$diet))

ggplot(auc_data, aes(x=RQ)) + 
  geom_density()

ylim <- max(auc_data$RQ)-3000
ggplot(auc_data, aes(x = diet, y = RQ)) +
  geom_boxplot(size = .75) +
  geom_jitter(alpha = .5) +
  stat_compare_means(aes(group=diet), label = "p.signif", method="t.test", label.y = ylim)+
  facet_grid(sex ~ interval, margins = TRUE) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))+
  labs(x = "", y = "RWL")

ggplot(auc_data, aes(x=interval, y=RQ, color = trt))+
  geom_boxplot()

#time intervals
#DAY
day <- filter(auc_data, interval == "light")

#The ANOVA
mod<-lm(RQ ~ diet * sex , day)
anova(mod)

#anova with x in covar as response
anovaX_mod<-lm(weight ~ diet * sex, day)
anova(anovaX_mod)

#transform
day$trans <- log10(10*day$RQ)
mod<-lm(trans ~ diet * sex, day)
anova(mod)

#TESTING ASSUMPTIONS
#testing for homo of slopes
slopes_mod<-lm(trans ~ sex + diet + weight + diet:weight, day)
anova(slopes_mod)

#Perform a Shapiro-Wilk test for normality of residuals
day$resids <- residuals(mod)
shapiro.test(day$resids)

#Perform Levene's Test for homogenity of variances
#library(car)
leveneTest(trans ~ diet, data = day)

#Perform a Tukey 1-df Test for Non-additivity
day$preds <- predict(mod)
day$sq_preds <- day$preds^2
tukey_mod<-lm(trans ~ sex + diet + sq_preds, day)
anova(tukey_mod)

#adjusting for differences in weight
ancova_mod<-lm(trans ~ diet + sex + weight, day)
Anova(ancova_mod, type = 2)
#to get beta val
summary(ancova_mod)

xbar <- mean(day$weight)
day$z <- day$trans + -0.003652*(day$weight - xbar)

anovaZ_mod<-lm(z ~ diet * sex, day)
anova(anovaZ_mod)

#post adj assumption testing
#norm of residuals 
day$anovaZ_resids <- residuals(anovaZ_mod)
shapiro.test(day$anovaZ_resids)

#homo of variety var
leveneTest(z ~ sex, data = day)

#Test for Nonadditivity
day$anovaZ_preds <- predict(anovaZ_mod)
day$sq_anovaZ_preds <- day$anovaZ_preds^2
tukeyZ_mod<-lm(z ~ sex + diet + sq_anovaZ_preds, day)
anova(tukeyZ_mod)

library(emmeans)
lsm <- emmeans(anovaZ_mod, specs = pairwise ~ sex:diet)
contrast(lsm, method = "pairwise", adjust = "tukey")
lsm$contrasts

#################
#NIGHT
night <- filter(auc_data, interval == "dark")
night$RQ <- as.numeric(night$RQ)

#The ANOVA
mod<-lm(RQ ~ diet * sex , night)
anova(mod)

#transform
night$trans <- log10(10*night$RQ)
mod<-lm(trans ~ diet * sex, night)
anova(mod)

#anova with x in covar as response
anovaX_mod<-lm(weight ~ diet * sex, night)
anova(anovaX_mod)

#TESTING ASSUMPTIONS
#testing for homo of slopes
#This interaction must be NS in order for us to carry out the ANCOVA.
slopes_mod<-lm(trans ~ sex + diet + weight + diet:weight, night)
anova(slopes_mod)

#failed without transformation...passed with transformation
#Perform a Shapiro-Wilk test for normality of residuals
night$resids <- residuals(mod)
shapiro.test(night$resids)

#Perform Levene's Test for homogenity of variances
#library(car)
leveneTest(trans ~ diet, data = night)

#Perform a Tukey 1-df Test for Non-additivity
night$preds <- predict(mod)
night$sq_preds <- night$preds^2
tukey_mod<-lm(trans ~ sex + diet + sq_preds, night)
anova(tukey_mod)


#adjusting for differences in weight
ancova_mod<-lm(trans ~ diet + sex + weight, night)
Anova(ancova_mod, type = 2)
#to get beta val
summary(ancova_mod)

xbar <- mean(night$weight)
night$z <- night$trans + 0.001153 *(night$weight - xbar)

anovaZ_mod<-lm(z ~ diet * sex, night)
anova(anovaZ_mod)

#post adj assumption testing
#norm of residuals 
night$anovaZ_resids <- residuals(anovaZ_mod)
shapiro.test(night$anovaZ_resids)

#homo of variety var
leveneTest(z ~ diet, data = night)

#Test for Nonadditivity
night$anovaZ_preds <- predict(anovaZ_mod)
night$sq_anovaZ_preds <- night$anovaZ_preds^2
tukeyZ_mod<-lm(z ~ sex + diet + sq_anovaZ_preds, night)
anova(tukeyZ_mod)


library(emmeans)
lsm <- emmeans(anovaZ_mod, specs = pairwise ~ sex:diet)
lsm$contrast

#################
#T1
t1 <- filter(auc_data, interval == "T1")
t1$RQ <- as.numeric(t1$RQ)

#The ANOVA
mod<-lm(RQ ~ diet * sex , t1)
anova(mod)

#transform
t1$trans <- log10(10*t1$RQ)
mod<-lm(trans ~ diet * sex, t1)
anova(mod)

#anova with x in covar as response
anovaX_mod<-lm(weight ~ diet * sex, t1)
anova(anovaX_mod)

#TESTING ASSUMPTIONS
#testing for homo of slopes
#This interaction must be NS in order for us to carry out the ANCOVA.
slopes_mod<-lm(trans ~ sex + diet + weight + diet:weight, t1)
anova(slopes_mod)

#Perform a Shapiro-Wilk test for normality of residuals
t1$resids <- residuals(mod)
shapiro.test(t1$resids)

#Perform Levene's Test for homogenity of variances
#library(car)
leveneTest(trans ~ diet, data = t1)

#Perform a Tukey 1-df Test for Non-additivity
t1$preds <- predict(mod)
t1$sq_preds <- t1$preds^2
tukey_mod<-lm(trans ~ sex + diet + sq_preds, t1)
anova(tukey_mod)

#adjusting for differences in weight
ancova_mod<-lm(trans ~ diet + sex + weight, t1)
Anova(ancova_mod, type = 2)
#to get beta val
summary(ancova_mod)

xbar <- mean(t1$weight)
t1$z <- t1$trans + -0.02070*(t1$weight - xbar)

anovaZ_mod<-lm(z ~ diet * sex, t1)
anova(anovaZ_mod)

#post adj assumption testing
#norm of residuals 
t1$anovaZ_resids <- residuals(anovaZ_mod)
shapiro.test(t1$anovaZ_resids)

#homo of variety var
leveneTest(z ~ sex, data = t1)

#Test for Nonadditivity
t1$anovaZ_preds <- predict(anovaZ_mod)
t1$sq_anovaZ_preds <- t1$anovaZ_preds^2
tukeyZ_mod<-lm(z ~ sex + diet + sq_anovaZ_preds, t1)
anova(tukeyZ_mod)

library(emmeans)
lsm <- emmeans(anovaZ_mod, specs = pairwise ~ sex:diet)
lsm$contrasts

#################
#T2
t2 <- filter(auc_data, interval == "T2")
t2$RQ <- as.numeric(t2$RQ)

#The ANOVA
mod<-lm(RQ ~ diet * sex , t2)
anova(mod)

#transform
t2$trans <- log10(10*t2$RQ)
mod<-lm(trans ~ diet * sex, t2)
anova(mod)

#anova with x in covar as response
anovaX_mod<-lm(weight ~ diet * sex, t2)
anova(anovaX_mod)

#TESTING ASSUMPTIONS
#testing for homo of slopes
#This interaction must be NS in order for us to carry out the ANCOVA.
slopes_mod<-lm(trans ~ sex + diet + weight + diet:weight, t2)
anova(slopes_mod)

#fails without transformations
#Perform a Shapiro-Wilk test for normality of residuals
t2$resids <- residuals(mod)
shapiro.test(t2$resids)

#Perform Levene's Test for homogenity of variances
#library(car)
leveneTest(trans ~ diet, data = t2)

#Perform a Tukey 1-df Test for Non-additivity
t2$preds <- predict(mod)
t2$sq_preds <- t2$preds^2
tukey_mod<-lm(trans ~ sex + diet + sq_preds, t2)
anova(tukey_mod)

#adjusting for differences in weight
ancova_mod<-lm(trans ~ diet + sex + weight, t2)
Anova(ancova_mod, type = 2)
#to get beta val
summary(ancova_mod)

xbar <- mean(t2$weight)
t2$z <- t2$trans + -0.005369*(t2$weight - xbar)

anovaZ_mod<-lm(z ~ diet * sex, t2)
anova(anovaZ_mod)

#post adj assumption testing
#norm of residuals 
t2$anovaZ_resids <- residuals(anovaZ_mod)
shapiro.test(t2$anovaZ_resids)

#homo of variety var
leveneTest(z ~ sex, data = t2)

#Test for Nonadditivity
t2$anovaZ_preds <- predict(anovaZ_mod)
t2$sq_anovaZ_preds <- t2$anovaZ_preds^2
tukeyZ_mod<-lm(z ~ sex + diet + sq_anovaZ_preds, t2)
anova(tukeyZ_mod)

library(emmeans)
lsm <- emmeans(anovaZ_mod, specs = pairwise ~ sex:diet)
lsm$contrasts


auc_master <- merge(auc_master,auc_data,by=c("Animal_ID","interval", "diet", "sex"))

auc_master <- auc_master %>% 
  rename(
    mgH2O = auc_data.mgH2O
    )

auc_master_final <- data.frame(auc_master$Animal_ID, auc_master$interval, auc_master$diet,auc_master$sex, auc_master$mgH2O,auc_master$EE, auc_master$RQ,auc_master$weight.x)
```

weight and electro t tests
```{r}
#electrolyte_data <- read.csv("~/Box Sync/Cactus_Mouse_Physiology/data/electrolyte_data.csv")

subset1 <- subset(electrolyte_data, electrolyte_data$experiment == "5015_lowfat")
subset2 <- subset(electrolyte_data, electrolyte_data$experiment == "baseline")

diet <- rbind(subset1, subset2)

diet$experiment <- as.factor(diet$experiment)
diet$sex <- as.factor(diet$sex)

diet_exp <- diet %>% select(1, 4)
diet_exp$experiment <- as.factor(diet_exp$experiment)
diet_exp$weight <- as.numeric(diet_exp$weight)
t.test(diet_exp$weight~diet_exp$experiment)

diet_sex <- diet %>% select(2, 4)
diet_sex$sex <- as.factor(diet_sex$sex)
diet_sex$weight <- as.numeric(diet_sex$weight)
t.test(diet_sex$weight~diet_sex$sex)


electros <- diet %>% select(1,5:ncol(diet))
lapply(electros[-1], function(x) t.test(x ~ electros$experiment))

electros <- diet %>% select(2,5:ncol(diet))
lapply(electros[-1], function(x) t.test(x ~ electros$sex))
```

```{r}
 ggplotRegression <- function (fit) {

    require(ggplot2)

    ggplot(fit$model, aes_string(x = names(fit$model)[2], y = names(fit$model)[1])) + 
      geom_point() +
      stat_smooth(method = "lm", col = "red") +
      labs(title = paste("Adj R2 = ",signif(summary(fit)$adj.r.squared, 5),
                         "Intercept =",signif(fit$coef[[1]],5 ),
                         " Slope =",signif(fit$coef[[2]], 5),
                         " P =",signif(summary(fit)$coef[2,4], 5)))
    }

#split by males, females, diet, response var, and light and dark phase for metabolic vars :( i hate this kind of science
#NA,BUN, hct, ica
#we would present the cool plots
#time for a for loop

regression_data <- select(auc_master, mgH2O, EE, RQ)

interval <- c("light","dark")
sex <- unique(auc_master$sex.x)
diet <- unique(auc_master$diet)
electros <- c("Na", "K", "Crea", "BUN", "Hct", "iCa")

plot_list <- list()
n=1
#females
  for (a in interval) 
    {
      for (d in diet) 
        {
          loop_data <- auc_master[ which(auc_master$interval==a & auc_master$sex.x=="F" & auc_master$diet==d), ]
          regression <- lm(EE ~ mgH2O, data = loop_data)
          print(a)
          print(d)
          print(summary(regression))
          plot1 <- ggplotRegression(regression)
          
          regression <- lm(EE ~ RQ, data = loop_data)
          print(a)
          print(d)
          print(summary(regression))
          plot2 <- ggplotRegression(regression)
          
          regression <- lm(mgH2O ~ RQ, data = loop_data)
          print(a)
          print(d)
          print(summary(regression))
          plot3 <- ggplotRegression(regression)
          plot_list[[n]] <- plot
          n=n+1
        }
  }

plot_list <- list()
n=1
for (a in interval) 
    {
  for (b in electros) 
    {
      for (d in diet) 
        {
          loop_data <- auc_master[ which(auc_master$interval==a & auc_master$sex.x=="F" & auc_master$diet==d), ]
          regression <- lm(as.formula(paste0(b, " ~ EE")), data = loop_data)
          print(a)
          print(d)
          print(b)
          print(summary(regression))
          plot1 <- ggplotRegression(regression)
          
          regression <- lm(as.formula(paste0(b, " ~ RQ")), data = loop_data)
          print(a)
          print(d)
          print(b)
          print(summary(regression))
          plot2 <- ggplotRegression(regression)
          
          regression <- lm(as.formula(paste0(b, " ~ mgH2O")), data = loop_data)
          print(a)
          print(d)
          print(b)
          print(summary(regression))
          plot2 <- ggplotRegression(regression)
          plot_list[[n]] <- plot
          n=n+1
        }
      }
    }

plot_list <- list()
n=1
#males
  for (a in interval) 
    {
      for (d in diet) 
        {
          loop_data <- auc_master[ which(auc_master$interval==a & auc_master$sex.x=="M" & auc_master$diet==d), ]
          regression <- lm(EE ~ mgH2O, data = loop_data)
          print(a)
          print(d)
          print(summary(regression))
          plot1 <- ggplotRegression(regression)
          
          regression <- lm(EE ~ RQ, data = loop_data)
          print(a)
          print(d)
          print(summary(regression))
          plot2 <- ggplotRegression(regression)
          
          regression <- lm(mgH2O ~ RQ, data = loop_data)
          print(a)
          print(d)
          print(summary(regression))
          plot3 <- ggplotRegression(regression)
          plot_list[[n]] <- plot
          n=n+1
        }
      }

plot_list <- list()
n=1
for (a in interval) 
    {
  for (b in electros) 
    {
      for (d in diet) 
        {
          loop_data <- auc_master[ which(auc_master$interval==a & auc_master$sex.x=="M" & auc_master$diet==d), ]
          regression <- lm(as.formula(paste0(b, " ~ EE")), data = loop_data)
          print(a)
          print(d)
          print(b)
          print(summary(regression))
          plot1 <- ggplotRegression(regression)
          
          regression <- lm(as.formula(paste0(b, " ~ RQ")), data = loop_data)
          print(a)
          print(d)
          print(b)
          print(summary(regression))
          plot2 <- ggplotRegression(regression)
          
          regression <- lm(as.formula(paste0(b, " ~ mgH2O")), data = loop_data)
          print(a)
          print(d)
          print(b)
          print(summary(regression))
          plot2 <- ggplotRegression(regression)
          plot_list[[n]] <- plot
          n=n+1
        }
      }
    }
multiplot(plotlist = plot_list, cols = 4)

```

example code for single regression and plot
regression <- lm(auc_master.EE ~ auc_master.mgH2O, data = auc_master_final[ which(auc_master_final$auc_master.interval=='dark'), ])
summary(regression)
p1 <- ggplotRegression(regression)


electrolyte box plots for diet 5015_modified (lowfat)
```{r}
plist_electro <- list()

all_electro_T <- melt(electrolyte_data_diet, id.vars = c("experiment","sex", "weight", "mouse_ID"), variable.name = "electro")
   
for (i in unique(all_electro_T$electro))
  {
  all_electro_s <- subset(all_electro_T, all_electro_T$electro == i)
  
  # Plot
  plist_electro[[i]] <-all_electro_s %>%
  ggplot(aes(x=electro, y=value, fill=diet)) +
    geom_violin() +
    #scale_fill_viridis(discrete = TRUE, alpha=0.6) +
    scale_fill_manual(values=c("#0571B0", "grey31", "#CA0020"))+
    geom_signif(comparisons = list(c("Standard", "Low Fat")),
                  map_signif_level = TRUE, textsize=3)+
    geom_jitter(color="black", size=0.4, alpha=0.9) +
    #theme_ipsum() +
    theme(legend.position = "none") +
    xlab("")+
    ylab("")+
    facet_wrap(~sex)
  
  }

library(gridExtra)
library(grid)

electo_plots <- gridExtra::grid.arrange(grobs = plist_electro)

save("diet_box_plot", 25, 15, electo_plots)
```

#### Electrolites
two-tailed t-test (t.test), adjusted for multiple comparisons (Bonferroni correction, p.adjust), to test for significant (p < 0.05) differences in response variables between sexes for electrolyes
```{r}
all_dependent_variables <- colnames(electrolyte_data_diet)
dependent_variables_e <- all_dependent_variables[c(5,6,7,8,9,11,12,13,14,15)]


BL_M_e <- subset(electrolyte_data_diet[electrolyte_data_diet$sex == 'M' & electrolyte_data_diet$experiment == "baseline",])
BL_F_e <- subset(electrolyte_data_diet[electrolyte_data_diet$sex == 'F' & electrolyte_data_diet$experiment == "baseline",])
LF_M_e <- subset(electrolyte_data_diet[electrolyte_data_diet$sex == 'M' & electrolyte_data_diet$experiment == "5015_lowfat",])
LF_F_e <- subset(electrolyte_data_diet[electrolyte_data_diet$sex == 'F' & electrolyte_data_diet$experiment == "5015_lowfat",])

### M and F (all)
MF_all_list = c("BL_M_e", "BL_F_e",
                "LF_M_e", "LF_F_e")
               
#Write header to mean/sd file
meanSD_header <- paste('DV', 'Dataset', 'mean', 'sd', sep=',')
write.table(meanSD_header, "mean_sd_eachTreatment_Xsex_electro.csv", sep=',', col.names = FALSE, row.names = FALSE, quote = FALSE)

#write head to t test results file
t_header <- paste('Data1', 'Data2', 'DV', 'p-value', 'unadj. p', sep=',')
write.table(t_header, "ttest_results_Xsex_electro.csv", sep=',', col.names = FALSE, row.names = FALSE, quote = FALSE)


#For each experiment
#Calcualte the average and SD of each dependent variable
count = 1
for(df in MF_all_list){
  print(df)
  for (DV in dependent_variables_e){
    print(DV)
    this_mean = mean(get(df)[[DV]])
    this_sd = sd(get(df)[[DV]])
    line = paste(DV, df, this_mean, this_sd, sep=',')
    print(line)
    print('\n')
    write.table(line, "mean_sd_eachTreatment_Xsex_electro.csv", sep = ',', append=TRUE, col.names = FALSE, row.names = FALSE, quote = FALSE)
    
    #two sample t test - test whether the mean differs from other exp
    templist <- list("BL_M_e", "BL_F_e",
                      "LF_M_e", "LF_F_e")
    
    # Adjust for multiple comparisons (Bonferonni)
    k <- length(templist)
    num_pw_comparisons <- k*(k-1) / 2 #where k is the number of conditions being compared
    for(data in templist){
      ts <- t.test(get(df)[[DV]], get(data)[[DV]], conf.level = 0.95)
      new_p <- p.adjust(ts$p.value, method = "bonferroni", num_pw_comparisons)
      this_line=paste(df, data, DV, new_p, ts$p.value, sep = ',')
      print(this_line)
      write.table(this_line, "ttest_results_Xsex_electro.csv", sep=',', append = TRUE, col.names = FALSE, row.names = FALSE, quote = FALSE)
    }
    count = count + 1
  }
}
```

